---
title: "CIS8392 - Group 4 - Flickr Classification Analysis"
author: Meghana Garaga | Manoj Kumar Potnuru | Aushutosh Mishra | Uma Shankari Tamalampudi
date: "12/06/2023"
output: html_document
---

## 1.Problem Statement
  
Design and implement a classification analysis system for images retrieved from Flickr, focusing on a specific set of categories.


## 2.URL Link for the data we collected using APIs 

We have used a Flickr API methods to collect the data that we wanted for our classification.The link is as follows
https://drive.google.com/drive/folders/1LYbF_gszFL7ZikmZ04gLUw_oUcT20VP6?usp=drive_link

## 3.A Data Summary

 id - Photo ids of the photos
 Descriptions - Description of the post/photo 
 Interest_Rate - The interest rate fixed by the bank for each customer
 faves - number of favorites or likes of a photo
 Comments - number of comments of a photo
 Views - Number of views of a photo
 Engagement_Score - Engagement score of each photo

```{r}
plot(cars)
```

## 4.Read the file

```{r}
install.packages("stringr")
library(stringr)
data<- read.csv("cars_ship_plane.csv", header = TRUE)
```

```{r}
# Install and load necessary libraries if not already installed
if (!require("tidyverse")) 
```


## 5.Preprocessing 

```{r}
# # Install the tidyverse package if not already installed
# if (!requireNamespace("tidyverse", quietly = TRUE)) {
#   install.packages("tidyverse")
# }

# Load the tidyverse package
library(tidyverse)

# # Your data
# data <- read.csv("your_file.csv")

# Remove rows with empty values in Caption, Photo_URL, and Tags columns
df_cleaned <- data %>%
  drop_na(Caption, Photo_URL, Tags)
# Print the cleaned data frame
print(df_cleaned)

```

## 6. Data Visualization

```{r}
# Install and load necessary libraries if not already installed
if (!require("tidyverse")) install.packages("tidyverse")

# Assuming 'data' is already loaded with your dataset

# Your existing code for extracting labels goes here

# Create a bar plot of label frequencies
label_counts <- table(df_cleaned$Label)
label_counts_df <- data.frame(Label = names(label_counts), Count = as.numeric(label_counts))

# Print label counts
print(label_counts_df)

# Plotting using ggplot2
library(ggplot2)
ggplot(label_counts_df, aes(x = Label, y = Count, fill = Label)) +
  geom_bar(stat = "identity") +
  labs(title = "Distribution of Labels",
       x = "Label",
       y = "Count") +
  theme_minimal()
```


## 7. 

```{r}
# Load the data from CSV file
data2 <- read.csv('cars_ship_plane.csv')

# Create a folder for images
output_folder_flickrimages <- "/CIS8398/flickerimages"
dir.create(output_folder_flickrimages, showWarnings = FALSE)

# Initialize a counter for the total number of images downloaded
total_images_downloaded <- 0

# Iterate through each URL and download the corresponding image
for (i in seq_along(data2$Photo_URL)) {
  url <- data2$Photo_URL[i]
  label <- data2$Label[i]
  photo_id <- data2$Photo_ID[i]

  # Create a subfolder for each label
  label_folder <- file.path(output_folder_flickrimages, label)
  dir.create(label_folder, showWarnings = FALSE)

  # Set the file path and name for the downloaded image
  file_path <- file.path(label_folder, paste0(photo_id, "_", label, ".jpg"))

  tryCatch({
    # Download the image and save it to the specified file path
    download.file(url, destfile = file_path, mode = "wb", quiet = TRUE)

    # Increment the total count of images downloaded
    total_images_downloaded <- total_images_downloaded + 1
  }, error = function(e) {
    # Print an error message if the download fails
    cat(sprintf("Error downloading image %d from URL %s: %s\n", i, url, e$message))
  })
}

# Print the final count of total images downloaded
cat("Total images downloaded:", total_images_downloaded, "\n")
```

```{r}
# Set your dataset paths on local desktop
car_dir <- "C:/CIS8398/flickerimages/car"
ship_dir <- "C:/CIS8398/flickerimages/ship"
plane_dir <- "C:/CIS8398/flickerimages/plane"

# Set your output paths for train and test on local desktop
train_dir <- "C:/CIS8398/flickerimages/train"
test_dir <- "C:/CIS8398/flickerimages/test"

# Function to split images into train and test
split_images <- function(vehicle_dir, vehicle_name) {
  # Get all image files in the vehicle folder
  image_files <- list.files(vehicle_dir, pattern = "\\.(jpg|jpeg|png)$", full.names = TRUE)

  # Randomly shuffle the image files
  set.seed(123)  # Set seed for reproducibility
  shuffled_files <- sample(image_files)

  # Determine the split point for train and test
  split_point <- round(0.8 * length(shuffled_files))

  # Create subfolders in train and test for the vehicle
  vehicle_train_dir <- file.path(train_dir, vehicle_name)
  vehicle_test_dir <- file.path(test_dir, vehicle_name)

  dir.create(vehicle_train_dir, showWarnings = FALSE, recursive = TRUE)
  dir.create(vehicle_test_dir, showWarnings = FALSE, recursive = TRUE)

  # Copy images to train folder
  file.copy(shuffled_files[1:split_point], vehicle_train_dir)

  # Copy images to test folder
  file.copy(shuffled_files[(split_point + 1):length(shuffled_files)], vehicle_test_dir)
}

# Split images for each vehicle using local paths
split_images(car_dir, "car")
split_images(ship_dir, "ship")
split_images(plane_dir, "plane")

# Print a message indicating successful split
cat("Images successfully split into train and test folders.\n")


```

```{r}
library(keras)
library(reticulate)
library(tidyverse)

# Set your dataset paths for car, ship, and plane
train_dir <- "C:/CIS8398/flickerimages/train"
test_dir <- "C:/CIS8398/flickerimages/test"

# Building your network
model_vehicle <- keras_model_sequential() %>%
  layer_conv_2d(filters = 32, kernel_size = c(3, 3), activation = "relu",
                input_shape = c(150, 150, 3)) %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 64, kernel_size = c(3, 3), activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_flatten() %>%
  layer_dense(units = 512, activation = "relu") %>%
  layer_dense(units = 3, activation = "softmax")  # 3 output units for car, ship, plane

model_vehicle %>% compile(
  optimizer = "adam",
  loss = "categorical_crossentropy",  # Use categorical crossentropy for multi-class classification
  metrics = c("accuracy")
)

# Data preprocessing
train_datagen <- image_data_generator(rescale = 1/255)
test_datagen <- image_data_generator(rescale = 1/255)

train_generator <- flow_images_from_directory(
  train_dir,
  train_datagen,
  target_size = c(150, 150),
  batch_size = 20,
  class_mode = "categorical",
  classes = c("car", "ship", "plane")  # Specify classes for multi-class
)

test_generator <- flow_images_from_directory(
  test_dir,
  test_datagen,
  target_size = c(150, 150),
  batch_size = 20,
  class_mode = "categorical",
  classes = c("car", "ship", "plane")  # Specify classes for multi-class
)

```

```{r}

# Fit the model to the data using the generator
history_vehicle <- model_vehicle %>%
  fit_generator(
    train_generator,
    steps_per_epoch = as.integer(10),  # Replace with steps per epoch
    epochs = 10,
    validation_data = test_generator,
    validation_steps = as.integer(10)  # Replace with validation steps
  )

# Plotting the model history
plot(history_vehicle)

```

```{r}
# Assuming you have already loaded your model and have test data
test_datagen <- image_data_generator(rescale = 1/255)
test_generator <- flow_images_from_directory(
  'C:/CIS8398/flickerimages/test',
  test_datagen,
  target_size = c(150, 150),
  batch_size = 20,
  class_mode = 'categorical'  # Change this to categorical for multi-class
)

# Evaluate the model on the test data
eval_result <- model_vehicle %>% evaluate(test_generator)

# Print the test loss
cat("Test accuracy:", eval_result[[2]], "\n")
```

```{r}
library(keras)
library(reticulate)
library(tidyverse)

# Set your dataset paths for car, ship, and plane
train_dir <- "C:/CIS8398/flickerimages/train"
test_dir <- "C:/CIS8398/flickerimages/test"

# Building your network
model_vehicle1 <- keras_model_sequential() %>%
  layer_conv_2d(filters = 32, kernel_size = c(3, 3), activation = "relu",
                input_shape = c(150, 150, 3)) %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 64, kernel_size = c(3, 3), activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_flatten() %>%
  layer_dense(units = 512, activation = "relu") %>%
  layer_dense(units = 3, activation = "softmax")  # 3 output units for car, ship, plane

model_vehicle1 %>% compile(
  optimizer = "adam",
  loss = "categorical_crossentropy",  # Use categorical crossentropy for multi-class classification
  metrics = c("accuracy")
)

# Data preprocessing
train_datagen1 <- image_data_generator(
  rescale = 1/255,
  rotation_range = 40,
  width_shift_range = 0.2,
  height_shift_range = 0.2,
  shear_range = 0.2,
  zoom_range = 0.2,
  horizontal_flip = TRUE,
  fill_mode = "nearest"
)

test_datagen1 <- image_data_generator(rescale = 1/255)

train_generator1 <- flow_images_from_directory(
  train_dir,
  train_datagen1,
  target_size = c(150, 150),
  batch_size = 20,
  class_mode = "categorical",
  classes = c("car", "ship", "plane")  # Specify classes for multi-class
)

test_generator1 <- flow_images_from_directory(
  test_dir,
  test_datagen1,
  target_size = c(150, 150),
  batch_size = 20,
  class_mode = "categorical",
  classes = c("car", "ship", "plane")  # Specify classes for multi-class
)

```

```{r}
# Fit the model to the data using the generator
history_vehicle1 <- model_vehicle1 %>%
  fit_generator(
    train_generator1,
    steps_per_epoch = as.integer(10),  # Replace with steps per epoch
    epochs = 10,
    validation_data = test_generator1,
    validation_steps = as.integer(10)  # Replace with validation steps
  )

# Plotting the model history
plot(history_vehicle1)
```

```{r}
# Assuming you have already loaded your model and have test data
test_datagen <- image_data_generator(rescale = 1/255)
test_generator <- flow_images_from_directory(
  'C:/CIS8398/flickerimages/test',
  test_datagen1,
  target_size = c(150, 150),
  batch_size = 30,
  class_mode = 'categorical'  # Change this to categorical for multi-class
)

# Evaluate the model on the test data
eval_result1 <- model_vehicle1 %>% evaluate(test_generator)

# Print the test loss
cat("Test accuracy:", eval_result1[[2]], "\n")
```
